Stability of Ensembles of Biological Networks

### 2014-09-11 journal.tex
add section on decomposition of stability computations in terms of strongly connected components

@@

>>>>

\def\tr{\mathrm{tr}}
\def\Path{\mathrm{Path}}
\def\hier{\mathrm{hier}}
\def\Vertex{\mathrm{Vertex}}

@@

\section{Decomposition of stability computations into strongly connected components}
$${\chi(M)'(z) \over \chi(M)(z)} = \tr (zI - M)^{-1}$$
where, $\chi(M)(z) = \det (zI - M)$ is the chararacteristic polynomial of
the matrix $M$ with $z$ a complex variable.  This formula is readily
verified when $M$ is diagonal and has the nice property of being
additive over decompositions of the matrix. Making the change of variables $x=\frac{1}{z}$, we have
$$\tr (I - xM)^n = \sum_{n=0}^\infty x^n \sum_{a \in \Vertex(G)} \sum_{p
\in \Path_n (G)(a,a)} F(p)$$
where $G$ is the graph of the matrix and $\Path_n (G;a,b)$ is the set of
paths of length $n$ from $a$ to $b$.  $(\Vertex(G), \Path(G), \cdot)$ is a category; $F$ is the functor from this category to the multiplicative semigroup of the real numbers which corresponds to the matrix $M$.

We now make use of the additivity.  Let $\hier$ be the functor which
sends $G$ to a poset $P$.  Then every element $p$ of $\Path(a,a)$ lies
in $\hier^{-1} (\hier(a))$.  Hence, our sum decomposes into partial sums
for the fibers of the points in the hierarchy which corresponds to a
factorization of the characteristic polynomial.
